---
title: "data_wrangling"
author: "Dave"
date: "10/22/2018"
output: html_document


```{r}
library(tidyverse)  #install.packages("tidyverse")
```

```{r}
## read gapminder csv. Note the readr:: prefix identifies which package it's in
gapminder <- readr::read_csv('https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv')
```
```{r}
gapminder
head(gapminder)
tail(gapminder)

str(gapminder)
summary(gapminder)

names(gapminder)
dim(gapminder)
c(nrow(gapminder), ncol(gapminder))

```{r}
filter(gapminder, lifeExp < 29)
filter(gapminder, country == "Mexico")
filter(gapminder, country %in% c("Mexico", "Peru", "Brazil"))
filter(gapminder, country == "Mexico", year == 2002)
filter(gapminder, country == "Sweden", lifeExp)
mean(lifeExp)
sweden <- filter(gapminder, country == "Sweden")

```

```
```{r}
select(gapminder, year, lifeExp)
select(gapminder, -continent, -lifeExp)
sweden <- filter(gapminder, country == "Sweden")
gap_cambodia <- filter(gapminder, country == "Cambodia")
gap_cambodia2 <- select(gap_cambodia, -continent, -lifeExp)
```{r}
gapminder %>% head()
```{r}

```

gap_cambodia <- gapminder %<%
filter(country == "Cambodia") %>%
select(-continent, -lifeExp)

```
```{r}
gap_cambodia <- gapminder %>% 
filter(country == "Cambodia") %>% 
select(-continent, -lifeExp)

gapminder %>% 
mutate(gdp = pop *gdpPercap)

gapminder %>% 
select(-continent, -lifeExp) %>% 
filter(country == "Egypt") %>% 
mutate(gdp = pop * gdpPercap) %>% 
mutate(max_gdp = max(gdp))

gapminder %>% 
select(-continent, -lifeExp) %>% 
filter(country == "Vietnam") %>% 
mutate(gdp   =pop*gdpPercap,
max_gdp = max(gdp)) 
```
```{r}
gapminder %>% 
group_by(country) %>%
mutate(gdp = pop * gdpPercap) %>%
summarize(max_gdp = max(gdp)) %>%
ungroup()


```
```{r}
gapminder %>%
  group_by(country) %>%
  mutate(gdp = pop * gdpPercap) %>%
  summarize(max_gdp = max(gdp)) %>%
  ungroup()
  
```
```{r}
co2 <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/co2.csv")
co2 %>% head ()
co2 %>% dim()
gap_2007 <- gapminder %>%
filter(year == 2007)
gap_2007 %>% dim ()
lj <- left_join(gap_2007, co2, by = "country")
lj %>% View()
lj %>% summary()
lj %>% dim()
lj <- gap_2007 %>% 
left_join(co2, by = "country")
rj <- right_join(gap_2007, co2, by = "country")
View(rj)





```



